---
title: "Stat 245 Tidying the Dataset"
author: "Eco Preserve Team: Trey, Hansol, Quan, and Hayworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
require(s245)
require(data.table)
library(dplyr)
library(lubridate)


# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


### Read in the Given Datasets for 2010-2016 Leaf Drop Data
```{r}
library(tidyverse)
## Skip lines and omit any blanks, NA's, or other insufficient cells
data2010 <- read.csv('2010 Leaf Drop Data Sheets.csv', na = c('', 'x', 'X'), skip = 3)
data2011 <- read.csv('2011 Leaf Drop Data Sheets.csv', na = c('', 'x', 'X'), skip = 3)
data2012 <- read.csv('2012 Leaf Drop Data Sheets.csv', na = c('', 'x', 'X'), skip = 3)
data2013 <- read.csv('2013 Leaf Drop Data Sheets.csv', na = c('', 'x', 'X'), skip = 4)
data2014 <- read.csv('2014 Leaf Drop Data Sheets.csv', na = c('', 'x', 'X', 'thanksgiving', 'Thanksgiving'), skip = 4)
data2015 <- read.csv('2015 Leaf Drop Data Sheets.csv', na = c('', 'x', 'X'), skip = 4)
data2016 <- read.csv('2016 Leaf Drop Data Sheets.csv', na = c('', 'x', 'X'), skip = 4)
```

### Rename Columns
```{r}
names(data2010)[1] <- 'Date'
names(data2010)[2] <- 'Basket'
names(data2010)[3] <- 'InitialWeight'
names(data2010)[4] <- 'After2Weeks'


names(data2011)[2] <- 'Date'
names(data2011)[3] <- 'Basket'
names(data2011)[4] <- 'InitialWeight'
names(data2011)[5] <- 'After2Weeks'


names(data2012)[2] <- 'Date'
names(data2012)[3] <- 'Basket'
names(data2012)[4] <- 'InitialWeight'
names(data2012)[5] <- 'After2Weeks'


names(data2013)[2] <- 'Date'
names(data2013)[3] <- 'Basket'
names(data2013)[4] <- 'InitialWeight'
names(data2013)[5] <- 'After2Weeks'


names(data2014)[2] <- 'Date'
names(data2014)[3] <- 'Basket'
names(data2014)[4] <- 'InitialWeight'
names(data2014)[5] <- 'After2Weeks'


names(data2015)[2] <- 'Date'
names(data2015)[3] <- 'Basket'
names(data2015)[4] <- 'InitialWeight'
names(data2015)[5] <- 'After2Weeks'


names(data2016)[2] <- 'Date'
names(data2016)[3] <- 'Basket'
names(data2016)[4] <- 'InitialWeight'
names(data2016)[5] <- 'After2Weeks'

## Keep only relevant columns
data2010 <- data2010 %>%
  select('Date', 'Basket', 'InitialWeight', 'After2Weeks')

data2011 <- data2011 %>%
  select('Date', 'Basket', 'InitialWeight', 'After2Weeks')

data2012 <- data2012 %>%
  select('Date', 'Basket', 'InitialWeight', 'After2Weeks')

data2013 <- data2013 %>%
  select('Date', 'Basket', 'InitialWeight', 'After2Weeks')

data2014 <- data2014 %>%
  select('Date', 'Basket', 'InitialWeight', 'After2Weeks')

data2015 <- data2015 %>%
  select('Date', 'Basket', 'InitialWeight', 'After2Weeks')

data2016 <- data2016 %>%
  select('Date', 'Basket', 'InitialWeight', 'After2Weeks')

```

### Pasting Together
```{r}
## Keep only necessary rows
data2010 <- data2010[1:75,]
data2011 <- data2011[1:114,]
data2012 <- data2012[1:105,]
data2013 <- data2013[1:133,]
data2014 <- data2014[1:129,]
data2015 <- data2015[1:115,]
data2016 <- data2016[1:124,]

## Bind the six datasets into one
newdata <- bind_rows(data2010, data2011)
newdata <- bind_rows(newdata, data2012)
newdata <- bind_rows(newdata, data2013)

## Make 2014 column for Basket an integer
data2014 <- data2014 %>%
  mutate(Basket = as.integer(Basket))


newdata <- bind_rows(newdata, data2014)
newdata <- bind_rows(newdata, data2015)

finaldata <- bind_rows(newdata, data2016)
```

### Replace dates labeled as '/16' to '/2016'
```{r}
## Rename informal dates
## Amend each informal date separately  so that days such as '10/16/2014' are not renamed to '10/2016/2014'
finaldata <- finaldata%>%
  rename(Date = 'Date')%>%
  mutate(Date = stringr::str_replace(Date, '11/23/16', '11/23/2016'))
finaldata <- finaldata%>%
  rename(Date = 'Date')%>%
  mutate(Date = stringr::str_replace(Date, '11/28/16', '11/28/2016'))
finaldata <- finaldata%>%
  rename(Date = 'Date')%>%
  mutate(Date = stringr::str_replace(Date, '11/29/16', '11/29/2016'))
finaldata <- finaldata%>%
  rename(Date = 'Date')%>%
  mutate(Date = stringr::str_replace(Date, '11/30/16', '11/30/2016'))
finaldata <- finaldata%>%
  rename(Date = 'Date')%>%
  mutate(Date = stringr::str_replace(Date, '12/01/16', '12/01/2016'))
finaldata <- finaldata%>%
  rename(Date = 'Date')%>%
  mutate(Date = stringr::str_replace(Date, '12/02/16', '12/02/2016'))
finaldata <- finaldata%>%
  rename(Date = 'Date')%>%
  mutate(Date = stringr::str_replace(Date, '12/03/16', '12/03/2016'))
finaldata <- finaldata%>%
  na.omit()
```

### Separate Date Variable into Month, Day, and Year Variables
```{r}
finaldata <- finaldata %>%
  separate(Date, into = c('Month', 'Day', 'Year'), sep = '/')

finaldata <- finaldata %>%
  mutate(Day = as.numeric(Day))
```

### Add New Date Variable and 1-365 Julian Date Variable
```{r}
finaldata$Date <- do.call(paste, list(finaldata$Month, finaldata$Day, finaldata$Year))
finaldata$Date <- as.Date(finaldata$Date, format=c("%m %d %Y"))

finaldata$julian_date <- as.numeric(format(finaldata$Date, "%j"))
```

### Add Week Variable
```{r}
finaldata <- finaldata %>%
  mutate(Week = week(Date))

finaldata <- finaldata %>%
  mutate(Week = as.numeric(Week))
```

### Add in Weather Data
```{r}
weather_data <- read.csv('https://sldr.netlify.app/data/gr-weather-data.csv') 
weather_data = subset(weather_data, select = -c(wdir, wpgt, tsun))
```

### Wrangle the Weather Data
```{r}
## Keep only columns of interest
weather_data <- weather_data %>%
  select(date, prcp, snow, wspd, tmax, tmin, tavg, wspd)

weather_data <- weather_data %>%
  rename(Date = date)
```

### Merging the Weather and Leaf Drop Data
```{r}
## Create Common Date Variable between both data sets
weather_data <- weather_data %>%
  separate(Date, into = c('Year', 'Month', 'Day'), sep = '-')

weather_data <- weather_data %>%
  mutate(Day = as.numeric(Day))

weather_data$Date <- do.call(paste, list(weather_data$Month, weather_data$Day, weather_data$Year))
weather_data$Date <- as.Date(weather_data$Date, format=c("%m %d %Y"))

## Merge by Date
merged_data <- merge(x = finaldata, y = weather_data, by = 'Date')
```

### Organize Final Dataset
```{r}
## Remove Repeated Columns
merged_data <- merged_data %>%
  select(-'Month.y', -'Day.y', -'Year.y')

## Rename for Better Understandability
merged_data <- merged_data %>%
  rename(Month = Month.x) %>%
  rename(Day = Day.x) %>%
  rename(Year = Year.x) %>%
  rename(Julian = julian_date) %>%
  rename(Precipitation = prcp) %>%
  rename(Snow = snow) %>%
  rename(WindSpeed = wspd) %>%
  rename(MaxTemperature = tmax) %>%
  rename(MinTemperature = tmin) %>% 
  rename(AverageTemperature = tavg)

glimpse(merged_data)

```
### North vs. South of Stream Variable 
```{r}
## Create Variable where Baskets 1-6, 11 are North of the Stream, and Baskets 7-10, 12 are South of the Stream
merged_data <- merged_data %>%
  mutate(NorthSouth = case_when(Basket <= 6 ~ 'N',
                                 Basket == 11 ~ 'N',
                                 Basket == 7 ~ 'S',
                                 Basket == 8 ~ 'S',
                                 Basket == 9 ~ 'S',
                                 Basket == 10 ~ 'S',
                                 Basket == 12 ~ 'S',))
```

### Conversion and Units
```{r}
## Convert Basket Leaf Mass to Leaf Mass per Square Meter
merged_data <- merged_data %>%
  mutate(After2Weeks = (After2Weeks/0.165))
```
Units for Variables:
After2Weeks (Leaf Drop Mass): Grams per Square Meter ($g/m^2$)
Precipitation: Millimeters ($mm$)
Snow: Millimeters ($mm$)
WindSpeed: Meters per Second ($m/s$)
Temperature: Degrees Celsius ($°C$)

### Write the File
```{r}
write.csv(merged_data,'leafdropdata.csv')
```


The "merged_data" after running everything is the Leaf Drop Data and Weather Data finalized.
It is written into a new csv file called 'leafdropdata'.




