---
title: "Stat 245 Graphics and Modeling"
author: "Eco Preserve Team: Trey Tipton, Hansol Kim, Quan Huynh, and Hayworth Anderson"
date: 'December 15, 2021'
output: 
  html_document:
    fig_height: 2.2
    fig_width: 4
  pdf_document:
    fig_height: 4
    fig_width: 8
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
library(viridis)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


### Questions We will Attempt to Answer

- How does the dry weight of leaves falling compare week to week?
- How do the seasons vary year to year? As in, does the leaf falling "season" begin earlier or later depending on the year?
- How does the weather affect the total dry weight of the leaves?

### Read in Completely Tidied Dataset
```{r}
leafdropdata <- read.csv('leafdropdata.csv')


head(leafdropdata)
```


### Create Average Weekly Leaf Drop Weight by Year Graph
```{r}
## Aggregate the mean for each week and put into new table
avg_weekly_data <- aggregate(leafdropdata$After2Weeks, list(leafdropdata$Year, leafdropdata$Week), FUN=mean)

## Rename variables
avg_weekly_data <- avg_weekly_data %>%
  rename(Year = 'Group.1') %>%
  rename(Week = 'Group.2') %>%
  rename(Mean_Week = 'x')

## Make Year Categorical
avg_weekly_data <- avg_weekly_data %>%
  mutate(Year = as.character(Year))

## Show first six rows of averaged data
head(avg_weekly_data)

## Create Line graph and facet by year
gf_line(Mean_Week ~ Week, color = ~Year, data = avg_weekly_data) %>%
  gf_labs(title ='Figure 1', x= 'Week of Year', y = 'Average Leaf Drop Mass (g/m²)', color = '') %>%
  gf_facet_wrap(~Year)
```
This graph shows the Average Weekly Leaf Drop Weight over time. The graph is faceted by year in order to compare and contrast the Leaf Drop Weight over time. It appears that each year, although slightly different from one another, all show a common trend of peaking at around Week 42-45 of the year, which is around the end of October. Interestingly, 2011 Leaf Drop weight has two distinct peaks, one at the beginning and one at the end of October. There could be several reasons for this, so we will have to take a look temperature in each year to see if 2011 has distinctly different weather that might cause the bimodal distribution. In 2013 there is also a different shape as the rise in leaf drop began earlier than the rest and the peak was much lower.

### Create Average Daily Leaf Drop Weight by Year Graph
```{r}
## Aggregate the mean for each day and put into new table
avg_daily_data <- aggregate(leafdropdata$After2Weeks, list(leafdropdata$Year, leafdropdata$Julian), FUN=mean)

## Rename variables
avg_daily_data <- avg_daily_data %>%
  rename(Year = Group.1) %>%
  rename(Day = Group.2) %>%
  rename(Mean_Day = x)

## Make Year Categorical
avg_daily_data <- avg_daily_data %>%
  mutate(Year = as.character(Year))

## Show first six rows of averaged data
head(avg_daily_data)

## Create Line graph and facet by year
gf_line(Mean_Day ~ Day, color = ~Year, data = avg_daily_data) %>%
  gf_labs(title ='Figure 2', x= 'Day of Year', y = 'Average Leaf Drop Mass (g/m²)') %>%
  gf_facet_wrap(~Year)
```
This graph of Average Daily Leaf Drop Weight over time gives us more insight into what is happening throughout each year. Again in 2011, there is two distinct peaks, whereas in the other years the peaks are not as distinct. Now we see a similar trend with a more specific value as the Leaf Drop Weights tend to peak around day 300 which is October 27th. 2016 has a slightly later peak and day 300 is actually in between 2011's peaks.


### Basket Graph
```{r}
gf_point(After2Weeks ~ Precipitation | Basket, data = leafdropdata)%>%
  gf_labs(title ='Figure 3', x= 'Precipitation (mm)', y = 'Leaf Drop Mass (g/m²)')
```
This graph shows a scatter plot of the Leaf Drop Weights by the Precipitation faceted by Basket number. As the Baskets were placed randomly around the Calvin Ecosystem Preserve, we are trying to see if any specific Basket happens to be placed in a spot where more leaves tend to fall. There does not appear to be any big trends, but some Baskets may collect slightly more leaves than others. Precipitation does not seem to have a huge effect on the leaf drop mass.

### Basket Graph 2
```{r, fig.height = 7, fig.width = 5}
gf_boxplot(After2Weeks ~ Basket, color = ~factor(Basket), data = leafdropdata) %>% 
  gf_facet_wrap(~Year, ncol = 1) %>% 
  gf_theme(axis.text.x = element_blank())%>%
  gf_labs(title ='Figure 4', x = '', y = 'Leaf Drop Mass (g/m²)', color = 'Basket')
```

### Temperature Over Time
```{r}
## Aggregate the mean weight and maximum temperature for each month and put into new table
avg_daily_data <- aggregate(leafdropdata$After2Weeks, list(leafdropdata$Year, leafdropdata$Julian), FUN=mean)
avg_daily_data2 <- aggregate(leafdropdata$MaxTemperature, list(leafdropdata$Year, leafdropdata$Julian), FUN=mean)

avg_daily_data <- merge(x = avg_daily_data, y = avg_daily_data2, by = c('Group.1', 'Group.2'))

## Rename variables
avg_daily_data <- avg_daily_data %>%
  rename(Year = Group.1) %>%
  rename(Day = Group.2) %>%
  rename(Mean_Daily_Temp = x.y) %>%
  rename(Mean_Daily_Weight = x.x)

## Make Year Categorical
avg_daily_data <- avg_daily_data %>%
  mutate(Year = as.character(Year))

## Create Line graph and facet by year
gf_line(Mean_Daily_Temp ~ Day, color = ~Mean_Daily_Weight, data = avg_daily_data) %>%
  gf_theme(scale_color_viridis_c())%>%
  gf_labs(title ='Figure 5', x= 'Day of Year', y = 'Temperature (°C)', color = 'Mean Daily Leaf Drop Weight') %>%
  gf_facet_wrap(~Year) 
  
```
This weather graph shows the Average Temperature each day throughout each year in Grand Rapids. We see what is expected as the temperature goes up and down somewhat randomly, but overall downward trends in temperature over time. In 2011, we see a drop in temperature that comes slightly earlier in the year than usual, between days 280 and 290. This could explain the earlier rise in Leaf Drop Weight that we saw earlier in the year 2011. After that peak, the temperature rose back up as the leaves fell less, and then the temperature went down again, hence the second peak. However, every year has inconsistent day to day weather, so it is difficult to attest Leaf Drop Weight to only Temperature and there are other predictors.

Overall, the Leaf Drop Weight goes up as temperature goes down, until it reaches the point where there are no more leaves to fall.

### Leaf Drop Weight by Average Temperature 
```{r}
gf_point(After2Weeks ~ AverageTemperature, color = ~Julian, data = leafdropdata) %>%
  gf_theme(scale_color_viridis_c())%>%
  gf_labs(title ='Figure 6', x= 'Average Temperature (ºC)', y = 'Leaf Drop Mass (g/m²)', color = 'Days')
```
This scatter plot shows the Leaf Drop Weight by the Minimum Temperature on the day that weight was recorded. The day is represented by color. We see a somewhat symmetric, unimodal, and normal distribution. The distribution shows us that there tends to be higher Leaf Drop weight as the temperature is in the middle, and lower Leaf Drop Weight if it is very hot or very cold. The color shows that the majority of high Leaf Drop Weights happen between around day 290-320. The darker and lighter points which resemble earlier and later dates in the year appear to be at the ends of the distribution, with low Leaf Drop Weights and either a high or low temperature.

### Leaf Drop Weight by Average Temperature 
```{r}
## Add Graph with Temperature by Leaf Drop by Year as Requested by Partner
gf_point(After2Weeks ~ AverageTemperature, data = leafdropdata) %>%
  gf_labs(title ='Figure 7', x= 'Average Temperature (°C)', y = 'Leaf Drop Mass (g/m²)', color = 'Year')%>%
  gf_facet_wrap(~Year) 
```
These scatter plots show that although Temperature varies year to year, there does not seem to be any trend and the temperature depending on the year is seemingly random.

### Leaf Drop Weights by Wind Speed by Month
```{r}
gf_point(After2Weeks ~ WindSpeed, data = leafdropdata) %>% 
  gf_labs(title ='Figure 8', x= 'Wind Speed (m/s)', y = 'Leaf Drop Mass (g/m²)')%>%
  gf_facet_wrap(~Month) 
```
This is the Leaf Drop as a function of Wind Speed faceted by month. October seems to have a negatively skewed normal distribution while November definitely seems to be uniformly distributed. This just goes to show that high winds before the Fall season do not necessarily make leaves fall.

### North vs. South of Stream 
```{r}
gf_boxplot(After2Weeks ~ NorthSouth, data = leafdropdata)%>%
  gf_labs(title ='Figure 9', x = "Day of Year", y = 'Leaf Drop Mass (g/m²)')
```
Baskets 1-6 and 11 lie North of the stream and 7-10 and 12 ar South of the Stream. There seems to be no difference between the Leaf Drop Mass on the North or South side of the stream.

### North vs. South of Stream Basket Graph
```{r}
## Create North vs. South of Stream scatter plots as Requested by Partner
gf_point(After2Weeks ~ Julian | NorthSouth, data = leafdropdata)%>%
  gf_labs(title ='Figure 10', x = "Day of Year", y = 'Leaf Drop Mass (g/m²)')
```
Both Leaf Drop Mass distributions over time look relatively similar for the baskets on the North and South side of the stream. They both have normal and unimodal distributions.

### Table of Leaf Drop Mass by Basket Across Years
```{r}
## Aggregate the average Leaf Drop Mass by Year and Basket
avg_basket_data <- aggregate(leafdropdata$After2Weeks, list(leafdropdata$Year, leafdropdata$Basket), FUN=mean)

## Rename Variables
avg_basket_data <- avg_basket_data %>%
  rename(Year = Group.1) %>%
  rename(Basket = Group.2) %>%
  rename(Mean_Basket = x)

## Create the Table as Requested by Partner
avg_basket_data %>%
  pivot_wider(names_from = Year, values_from = Mean_Basket)
```

### Questions We Answered
- How does the dry weight of leaves falling compare week to week?

The Leaf Drop Mass increases over the Fall season, hits a peak, then trends downward. The upward trend is likely due to the days getting shorter and weather getting colder, whereas the downward trend is likely due to a decrease in the possible amount of leafs to fall.

- How do the seasons vary year to year? As in, does the leaf falling "season" begin earlier or later depending on the year?

The Year to Year Leaf Drop masses are relatively similar as the Fall season continues. However, there are a few abnormalities: 2011 has two peaks and 2013 has an early increase in leaf drop leading to less of a peak.

- How does the weather affect the total dry weight of the leaves?

It seems that weather is less important than time. Obviously, as it gets colder, the Leaf Drop Mass increases and then there are less leaves to fall so it decreases.
  